---
title: "Quantitative Investment Analysis Homework 2"
author: "Refik Türkeli"
date: "30/11/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

### We would like to determine if the typical amount spent per customer for lunch at a new restaurant in Istanbul is more than TL60.00. A sample of 45 customers over a ten-week period was randomly selected and the average amount spent was TL45.50. Assume that the standard deviation is known to be TL6.50. Using a 0.05 level of significance, can we conclude the typical amount spent per customer is more than TL60.00?

```{r}
null_value <- 60
sample_n <- 45
sample_mean <- 45.5
stdev <- 6.5
a <- 0.05
```

### Answer

$$
\begin{aligned}
H_0 : \mu &\le `r null_value`\\
H_a : \mu &> `r null_value`
\end{aligned}
$$

```{r}
standard_error <- stdev / sqrt(sample_n)
rejection_point <- qt(a, sample_n - 1, lower.tail = FALSE)
test_statistic <- (sample_mean - null_value) / standard_error
```

$$
\begin{aligned}
\sigma_{\bar{X}} &= `r round(standard_error, 2)`\\
\text{Test Statistic} &= `r round(test_statistic, 2)`\\
\text{Rejection Point} &= `r round(rejection_point, 2)`\\
\end{aligned}
$$

No we can't conclude an increase. Test statistic is less than the rejection point.

\newpage

## Question 2

### Suppose an equity research analyst claims that the mean time to write a research report is at most 4 weeks. A sample of 14 analysts is randomly selected and it is found that the mean time taken by them to write a report was 2.5. Assume also that the standard deviation is known to be 0.6 weeks. Assuming the time to write a research report is normally distributed and using a 0.02 level of significance, would you conclude the analyst’s claim is true?

```{r}
null_value <- 4
sample_n <- 14
sample_mean <- 2.5
stdev <- 0.6
a <- 0.02
```

### Answer

$$
\begin{aligned}
H_0 : \mu &> `r null_value`\\
H_a : \mu &\le `r null_value`
\end{aligned}
$$

```{r}
standard_error <- stdev / sqrt(sample_n)
rejection_point <- qnorm(a)
test_statistic <- (sample_mean - null_value) / standard_error
```

$$
\begin{aligned}
\sigma_{\bar{X}} &= `r round(standard_error, 2)`\\
\text{Test Statistic} &= `r round(test_statistic, 2)`\\
\text{Rejection Point} &= `r round(rejection_point, 2)`\\
\end{aligned}
$$
Yes we can conlude that the analyst's claim is true. Test statistic is less than the rejection point.

\newpage

## Question 3

### According to a 2010 demographic report, the average Turkish. household spends TL40 per day. Suppose you recently took a random sample of 30 households in Yalova and the results revealed a mean of TL34.50. Suppose the standard deviation is known to be TL6.50. Using a 0.05 level of significance, can it be concluded that the average amount spent per day by Turkish households has decreased?

```{r}
null_value <- 40
sample_n <- 30
sample_mean <- 34.5
stdev <- 6.5
a <- 0.05
```

### Answer

$$
\begin{aligned}
H_0 : \mu &\ge `r null_value`\\
H_a : \mu &< `r null_value`
\end{aligned}
$$

```{r}
standard_error <- stdev / sqrt(sample_n)
rejection_point <- qt(a, df = sample_n - 1)
test_statistic <- (sample_mean - null_value) / standard_error
```

$$
\begin{aligned}
\sigma_{\bar{X}} &= `r round(standard_error, 2)`\\
\text{Test Statistic} &= `r round(test_statistic, 2)`\\
\text{Rejection Point} &= `r round(rejection_point, 2)`\\
\end{aligned}
$$

Yes we can conclude that average spending has decreased. Test statistic is less than the rejection point.

\newpage

## Question 4

### Historically, evening long-distance calls from a particular city have averaged 17.2 minutes per call. In a random sample of 35 calls, the sample mean time was 15.8 minutes. Assume the standard deviation is known to be 4 minutes. Using a 0.10 level of significance, is there sufficient evidence to conclude that the average evening long-distance call has decreased?


```{r}
null_value <- 17.2
sample_n <- 35
sample_mean <- 15.8
stdev <- 4
a <- 0.10
```

### Answer

$$
\begin{aligned}
H_0 : \mu &\ge `r null_value`\\
H_a : \mu &< `r null_value`
\end{aligned}
$$

```{r}
standard_error <- stdev / sqrt(sample_n)
rejection_point <- qt(a, df = sample_n - 1)
test_statistic <- (sample_mean - null_value) / standard_error
```

$$
\begin{aligned}
\sigma_{\bar{X}} &= `r round(standard_error, 2)`\\
\text{Test Statistic} &= `r round(test_statistic, 2)`\\
\text{Rejection Point} &= `r round(rejection_point, 2)`\\
\end{aligned}
$$

There is enough evidence to conclude that long-distace call has decreased. Test statistic is less than the rejection point.

\newpage

## Question 5

### Suppose a food production line operates with a mean filling weight of 15 kilograms per container. Since over- or under-filling can be dangerous, a quality control inspector samples 30 items to determine whether or not the filling weight has to be adjusted. The sample revealed a mean of 15.32 kilograms. From past data, the standard deviation is known to be 0.75 kilograms. Using a 0.05 level of significance, can it be concluded that the process is out of control (not equal to 15 kilograms)?

```{r}
null_value <- 15
sample_n <- 30
sample_mean <- 15.32
stdev <- 0.75
a <- 0.05
```

### Answer

$$
\begin{aligned}
H_0 : \mu &= `r null_value`\\
H_a : \mu &\neq `r null_value`
\end{aligned}
$$

```{r}
standard_error <- stdev / sqrt(sample_n)
rejection_point <- qt(a / 2, df = sample_n - 1, lower.tail = FALSE)
test_statistic <- (sample_mean - null_value) / standard_error
```

$$
\begin{aligned}
\sigma_{\bar{X}} &= `r round(standard_error, 2)`\\
\text{Test Statistic} &= `r round(test_statistic, 2)`\\
\text{Rejection Point} &= `r round(rejection_point, 2)`\\
\end{aligned}
$$
We can conclude that process is out of control on the heavier side. Test statistic is higher than rejection point.